---
title: "Exercise1.01"
author: "Pradeep"
date: "12/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Read in the data
This section will read in the data. And print a summary

```{r}
bikesharing <- read_csv("./Data/day.csv")
head(bikesharing)
```

## Initial transformations for some of the data to make
The following are the mapping according to the data dictionary.
season_mapping = {1:"winter", 2:"spring", 3:"summer", 4:"fall"}
yr_mapping = {0: "2011", 1:"2012"}
mnth_mapping = {1:"Jan",2:"Feb",3:"Mar",4:"Apr",5:"May",6:"Jun",7:"Jul",8:"Aug",9:"Sep",10:"Oct",11:"Nov",12:"Dec"}
weekday_mapping={0: 'Sunday', 1: 'Monday', 2: 'Tuesday',3: 'Wednesday', 4: 'Thursday', 5: 'Friday', 6: 'Saturday'}
weather_mapping={1:"Clear",2:"Misty",3:"Light Snow/Rain",4:"Heavy Snow/Rain"}

```{r}
library(tidyverse)
library(magrittr)

bikesharing <- read_csv("./Data/day.csv")
# yr_mapping = {0: "2011", 1:"2012"}
# mnth_mapping = {1:"Jan",2:"Feb",3:"Mar",4:"Apr",5:"May",6:"Jun",7:"Jul",8:"Aug",9:"Sep",10:"Oct",11:"Nov",12:"Dec"}
# weekday_mapping={0: 'Sunday', 1: 'Monday', 2: 'Tuesday',3: 'Wednesday', 4: 'Thursday', 5: 'Friday', 6: 'Saturday'}
# weather_mapping={1:"Clear",2:"Misty",3:"Light Snow/Rain",4:"Heavy Snow/Rain"}

bikesharing %<>% mutate(
                        season=case_when(
                          season==1 ~ "winter",
                          season==2 ~ "spring",
                          season==3 ~ "summer",
                          season==4 ~ "fall",
                          TRUE~"NA"  ), 
                        yr = case_when(
                          yr==0 ~"2011",
                          yr==1 ~ "2012"
                        ),
                        mnth = case_when(
                          mnth==1 ~"Jan",
                          mnth==2 ~"Feb",
                          mnth==3 ~"Mar",
                          mnth==4 ~"Apr",
                          mnth==5 ~"May",
                          mnth==6 ~"Jun",
                          mnth==7 ~"Jul",
                          mnth==8 ~"Aug",
                          mnth==9 ~"Sep",
                          mnth==10 ~"Oct",
                          mnth==11 ~"Nov",
                          mnth==12 ~"Dec",
                          ),
                        weekday= case_when(
                          weekday==0 ~ "Sunday",
                          weekday==1 ~ "Monday",
                          weekday==2 ~ "Tuesday",
                          weekday==3 ~ "Wednesday",
                          weekday==4 ~ "Thursday",
                          weekday==5 ~ "Friday",
                          weekday==6 ~ "Saturday"
                        ),
                        weathersit= case_when(
                          weathersit==1 ~"Clear",
                          weathersit==2 ~"Misty",
                          weathersit==3 ~"Light Snow or Rain",
                          weathersit==4 ~"Heavy Snow or Rain"
                        )
                        
                          )
bikesharing %>% slice_sample(n=200)
```

