---
title: "Exercise1.01"
author: "Pradeep"
date: "12/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(zoo)
```
## Read in the data
This section will read in the data. And print a summary

## Initial transformations for some of the data to make
The following are the mapping according to the data dictionary.
season_mapping = {1:"winter", 2:"spring", 3:"summer", 4:"fall"}
yr_mapping = {0: "2011", 1:"2012"}
mnth_mapping = {1:"Jan",2:"Feb",3:"Mar",4:"Apr",5:"May",6:"Jun",7:"Jul",8:"Aug",9:"Sep",10:"Oct",11:"Nov",12:"Dec"}
weekday_mapping={0: 'Sunday', 1: 'Monday', 2: 'Tuesday',3: 'Wednesday', 4: 'Thursday', 5: 'Friday', 6: 'Saturday'}
weather_mapping={1:"Clear",2:"Misty",3:"Light Snow/Rain",4:"Heavy Snow/Rain"}

```{r}
library(tidyverse)
library(magrittr)

bikesharing <- read_csv("./Data/hour.csv")
# yr_mapping = {0: "2011", 1:"2012"}
# mnth_mapping = {1:"Jan",2:"Feb",3:"Mar",4:"Apr",5:"May",6:"Jun",7:"Jul",8:"Aug",9:"Sep",10:"Oct",11:"Nov",12:"Dec"}
# weekday_mapping={0: 'Sunday', 1: 'Monday', 2: 'Tuesday',3: 'Wednesday', 4: 'Thursday', 5: 'Friday', 6: 'Saturday'}
# weather_mapping={1:"Clear",2:"Misty",3:"Light Snow/Rain",4:"Heavy Snow/Rain"}

bikesharing %<>% mutate(
                        season=case_when(
                          season==1 ~ "winter",
                          season==2 ~ "spring",
                          season==3 ~ "summer",
                          season==4 ~ "fall",
                          TRUE~"NA"  ), 
                        yr = case_when(
                          yr==0 ~"2011",
                          yr==1 ~ "2012"
                        ),
                        mnth = case_when(
                          mnth==1 ~"Jan",
                          mnth==2 ~"Feb",
                          mnth==3 ~"Mar",
                          mnth==4 ~"Apr",
                          mnth==5 ~"May",
                          mnth==6 ~"Jun",
                          mnth==7 ~"Jul",
                          mnth==8 ~"Aug",
                          mnth==9 ~"Sep",
                          mnth==10 ~"Oct",
                          mnth==11 ~"Nov",
                          mnth==12 ~"Dec",
                          ),
                        weekday= case_when(
                          weekday==0 ~ "Sunday",
                          weekday==1 ~ "Monday",
                          weekday==2 ~ "Tuesday",
                          weekday==3 ~ "Wednesday",
                          weekday==4 ~ "Thursday",
                          weekday==5 ~ "Friday",
                          weekday==6 ~ "Saturday"
                        ),
                        weathersit= case_when(
                          weathersit==1 ~"Clear",
                          weathersit==2 ~"Misty",
                          weathersit==3 ~"Light Snow or Rain",
                          weathersit==4 ~"Heavy Snow or Rain"
                        )
                        
                          )
bikesharing %>% slice_sample(n=200)
```

We will plot the distribution plot of the registered and casual riders all up

```{r}
bikeshare_plot<-bikesharing %>% pivot_longer(cols = c(registered,casual),names_to="RideType",values_to="NumRides") %>% select(RideType,NumRides)
fig <- ggplot(data = bikeshare_plot) +
  geom_density(mapping = aes(x=NumRides,fill=RideType),alpha=0.25) +
  scale_y_continuous(labels=function(n){format(n, scientific = FALSE)})+
  theme_classic() +
  theme(
    legend.position = c(.95, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6)
    )
fig
ggsave("./figs/rides_distributions.png",plot = fig,device = "png")
```
Plot the time series. Evolution of Rides by time.

```{r}

plot_data<- bikesharing %>% select(dteday,registered,casual) %>% 
  group_by(dteday) %>% summarise(registered=sum(registered),casual=sum(casual)) %>% 
  pivot_longer(cols=c(registered,casual),names_to="RideType",values_to="NumberOfRides")

fig<-ggplot(data = plot_data) + 
  geom_line(mapping = aes(x=dteday,y=NumberOfRides,color=RideType))+
  theme_classic()+
  theme(
    legend.position = c(1, 1),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 0, 0, 0)
  )
fig
ggsave("./figs/rides_daily.png",plot = fig,device = "png")

```
Smoothed Plotting of Time Series
```{r}
windowsize <- 7
plot_data<- bikesharing %>% select(dteday,registered,casual) %>% 
  group_by(dteday) %>% 
  summarise(sumreg=sum(registered),sumcas=sum(casual)) %>% 
  mutate(regrollm=rollmean(x=sumreg,k=windowsize,align="right",fill=NA),
         casrollm=rollmean(x=sumcas,k=windowsize,align="right",fill=NA),
         regrollsd=rollapplyr(data= sumreg,width=windowsize,FUN="sd",fill=NA),
         casrollsd=rollapplyr(data= sumcas,width=windowsize,FUN="sd",fill=NA)
           
         ) %>% select(dteday,registered=regrollm,registeredsd=regrollsd,casual=casrollm,casualsd=casrollsd) %>% filter(!is.na(registered)) 
plot_data_1<- plot_data %>% select(dteday,registered,casual) %>% 
  pivot_longer(cols = c("registered","casual"),names_to="RideType",values_to="RollingMeanRides")
plot_data_2<- plot_data %>% select(dteday,registeredsd,casualsd) %>% 
  pivot_longer(cols = c("registeredsd","casualsd"),names_to="RideType",values_to="RollingsdRides")

fig <- ggplot(data = plot_data_1) + geom_line(mapping = aes(x=dteday,y=RollingMeanRides,color=RideType)) +
  geom_ribbon(data=plot_data_2,mapping = aes(x=dteday,ymin=plot_data_1$RollingMeanRides-2*RollingsdRides,ymax=plot_data_1$RollingMeanRides+2*RollingsdRides,fill=RideType),alpha=0.25,show.legend = FALSE) +
  theme_classic()+
  theme(
    legend.position = c(1, 1),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 0, 0, 0)
  )
fig
ggsave("./figs/rolling_rides_daily.png",plot = fig,device = "png")

 

# fig<-ggplot(data = plot_data) + 
#   geom_line(mapping = aes(x=dteday,y=NumberOfRides,color=RideType))+
#   theme_classic()+
#   theme(
#     legend.position = c(1, 1),
#     legend.justification = c("right", "top"),
#     legend.box.just = "right",
#     legend.margin = margin(0, 0, 0, 0)
#   )
# fig
# ggsave("./figs/rides_daily.png",plot = fig,device = "png")

```
# distribution of bike rides during a day and then during days of a week
